An Efficient Minibatch Acceptance Test for Metropolis-Hastings

### 2017-03-30 uai2017/paper.tex
fix organization

\section{EXPERIMENTS}\label{sec:experiments}
In one of the most closely-related prior
works,~\citep{cutting_mh_2014}, asymptotic CLT arguments are used to
argue that a modified standard M-H test can be used on subsets of data
with modest error. This assumes knowledge of dataset-wide mean
$\mu_{\rm std}$ each iteration (it depends on $\theta$). Determining
$\mu_{\rm std}$ exactly requires a scan over the entire dataset (so
the method in practice provides no data reduction), or some
model-specific bounds must be derived.  \citep{cutting_mh_2014} also
propose a conservative variant which assumes $\mu_{\rm std}=0$ and
avoids the scan.  We refer to the conservative version as {\sc
  AustereMH(c)} and the non-conservative variant as {\sc
  AustereMH(nc)}. We analyze both in the sequel.

In another related work~\citep{icml2014c1_bardenet14} use concentration bounds
and a similar modification to the standard M-H test ({\sc MHSubLhd} method). 
For {\sc MHSubLhd}, the test requires direct bounds on the log likelihood ratio,
which for general distributions requires knowing $p(x_i | \theta)$ and $p(x_i |
\theta')$ for all $N$ samples. In~\citep{icml2014c1_bardenet14}, explicit bounds
are given and depend on
\begin{equation}\label{eq:bad_bound}
    C_{\theta, \theta'} = \max_{1\leq i\leq N}|\log p(x_i | \theta') - \log p(x_i | \theta)|.
\end{equation}
It is possible to analytically determine $C_{\theta,\theta'}$ for logistic
regression and other simple models, but it is difficult to do so for more
complicated models, so one generally needs to use all $p(x_i | \theta')$
terms\footnote{The sample code provided by~\citep{icml2014c1_bardenet14}
computes $C_{\theta,\theta'}$ by traversing the entire data, thus providing no
performance advantage over the complete test.} to get $C_{\theta,\theta'}$, like
the {\sc AustereMH(nc)} test.  In contrast, we use quantitative forms of the CLT
which rely on measurable statistics from a \emph{single} minibatch.

@@

% Daniel: NOTE, I'm calling temperature "K" so that "T" stays reserved for
% "number of samples of \theta we draw".
In Sections~\ref{ssec:gaussians} and~\ref{ssec:logistic}, we benchmark {\sc
MHminibatch} against the tests from~\citep{icml2014c1_bardenet14}
and~\citep{cutting_mh_2014}, with the latter using a grid-search over minibatch
sizes $m$ and per-test thresholds $\epsilon$ described in
Appendix~\ref{sssec:grid_search}. Throughout our descriptions, we refer to a
\emph{trial} as the period when an algorithm collects all its desired samples
$\{\theta_1,\ldots,\theta_T\}$, generally with $T=3000$ or $T=5000$.

>>>>

\section{EXPERIMENTS}\label{sec:experiments}
Here we compare with the most similar prior works
~\citep{cutting_mh_2014} and~\cite{icml2014c1_bardenet14}.
In~\citep{cutting_mh_2014}, an asymptotic CLT is used to argue that a
modified standard M-H test can be used on subsets of the data. This
assumes knowledge of dataset-wide mean $\mu_{\rm std}$ each iteration
(it depends on $\theta$). Determining $\mu_{\rm std}$ exactly requires
a scan over the entire dataset, or some model-specific bounds.
\citep{cutting_mh_2014} also propose a conservative variant which
assumes $\mu_{\rm std}=0$ and avoids the scan.  We refer to the
conservative version as {\sc AustereMH(c)} and the non-conservative
variant as {\sc AustereMH(nc)}. We analyze both in this section.

In~\citep{icml2014c1_bardenet14}
concentration bounds are used with a similar modification to the standard M-H
test ({\sc MHSubLhd} method).  For {\sc MHSubLhd}, the required global
bound is denoted $C_{\theta, \theta'}$ which once again depends on
$\theta$ and so must be recomputed at each step, or estimated in a
model-specific way.   We obtained
sample code for both methods from the authors, and found that both
{\sc AustereMN(nc)} and {\sc MHSubLhd} scanned the entire dataset at
each iteration to derive these bounds. We do not include the cost of
doing this in our experiments, since otherwise there would be no
improvement over testing the full dataset. However, it should be kept
in mind that such bounds must be provided to these methods.
Our test by contrast uses a quantitative form of the CLT
which rely on measurable statistics from a \emph{single} minibatch.
It therefore requires no dataset-wide scans, and can be used, e.g. on
streams of data.

@@

% Daniel: NOTE, I'm calling temperature "K" so that "T" stays reserved for
% "number of samples of \theta we draw".
In Sections~\ref{ssec:gaussians} and~\ref{ssec:logistic}, we benchmark {\sc
MHminibatch} against {\sc MHSubLhd}, {\sc AustereMH(c)} and {\sc AustereMH(nc)}.
Hyperparameters for the latter were optimized using a grid-search over minibatch
sizes $m$ and per-test thresholds $\epsilon$ described in
Appendix~\ref{sssec:grid_search}. Throughout our descriptions, we refer to a
\emph{trial} as the period when an algorithm collects all its desired samples
$\{\theta_1,\ldots,\theta_T\}$, generally with $T=3000$ or $T=5000$.
